---
title: "Beta-carotene CHD triangulation example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Beta-carotene CHD triangulation example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(dplyr)
library(tidyr)
library(readxl)
library(tibble)
library(magrittr)
library(metafor)
library(triangulate)

```

## Create CHD dataset

We first define observational dietary studies from Aune et al.

```{r create data}
##########################
### Creating datapoint ###
##########################
# For observational studies directly from Aune's paper Figure 8
# For RCT studies, directly from Supplementary Figure 18 for total CHD risk and Supplementary Figure 22 for CHD mortality risk

### Diet ###
CHD_diet<-data.frame(Author=c("Osganian", "Klipstein-Grobusch", "Todd", "Pandey"),
                     RR=c(0.83, 0.22, 0.87, 0.84),
                     RR_LC=c(0.72, 0.07, 0.68, 0.66),
                     RR_UC=c(0.96, 0.68, 1.12, 1.09),
                     Study=rep("Obs", 4), Design=rep("Diet", 4), Dose=rep(1, 4))

### Circulatory ###
#Aune's results is per 25 ug/dL
sd_circ<-sqrt(78)*(((0.48-0.46)/0.01863)/3.92)
sd_diet<-sqrt(67)*((4465.5-3383.8)/3.92)
conv_factor<-0.27*(sd_diet/sd_circ)

CHD_circ<-data.frame(Author=c("Karppi", "Koh", "Hak"),
                     RR=c(0.64, 0.95, 0.82),
                     RR_LC=c(0.45, 0.66, 0.63),
                     RR_UC=c(0.93, 1.37, 1.07),
                     Study=rep("Obs", 3), Design=rep("Circ", 3), Dose=rep(25*conv_factor, 3)) %>%
  mutate(RR=exp((5000/Dose)*log(RR)), RR_LC=exp((5000/Dose)*log(RR_LC)),
         RR_UC=exp((5000/Dose)*log(RR_UC)), Dose=Dose^0) # convert RR to per 5000 ug/d

### RCT ###
CHD_rct<-data.frame(Author=c("Tornwall", "Cook", "Hennekens"),
                    RR=c(1.03, 1.00, 0.96),
                    RR_LC=c(0.92, 0.89, 0.85),
                    RR_UC=c(1.16, 1.13, 1.08),
                    Study=rep("RCT", 3), Design=rep("RCT", 3), Dose=c(20000, 25000, 25000)) %>%
  mutate(RR=exp(log(RR)/(Dose/5000)), RR_LC=exp(log(RR_LC)/(Dose/5000)),
         RR_UC=exp(log(RR_UC)/(Dose/5000)), Dose=Dose^0) # convert RR to per 5000 ug/d

### MR ###
# the conversion of per log transformed to per 1 unit of beta-carotene
# is based on Rodr?guez_Barranco2017
# conv_factor is based on µmol/L not µg/L
sd_circ<-sqrt(78)*((((0.48-0.46)*10)/0.01863)/3.92)
sd_diet<-sqrt(67)*((4465.5-3383.8)/3.92)
conv_factor<-0.27*(sd_diet/sd_circ)

CHD_MR<-data.frame(Author=c("CARDIoGRAMplusC4D", "UKBB", "FinnGen"),
                   RR=c(exp((5000/conv_factor)*(log(1.054)/301)),
                        exp((5000/conv_factor)*(log(1.015)/301)),
                        exp((5000/conv_factor)*(log(1.027)/301))),
                   RR_LC=c(exp((5000/conv_factor)*(log(0.948)/301)),
                           exp((5000/conv_factor)*(log(0.925)/301)),
                           exp((5000/conv_factor)*(log(0.840)/301))),
                   RR_UC=c(exp((5000/conv_factor)*(log(1.173)/301)),
                           exp((5000/conv_factor)*(log(1.114)/301)),
                           exp((5000/conv_factor)*(log(1.255)/301))),
                   Study=rep("MR", 3), Design=rep("Circ", 3), Dose=rep(1, 3))


### Combine all ###
Metadta_all_CHD<-CHD_diet %>%
  rbind(CHD_circ,CHD_rct, CHD_MR) %>%
  mutate(logRR=log(RR), se_logRR=(log(RR_UC)-log(RR_LC))/3.92, id=1:13, Outcome="CHD") %>%
  select(Author,Study, id,Dose, logRR, se_logRR, Design, Outcome)
```

## Merge with Risk of Bias (RoB) judgements

We merge effect estimates with risk-of-bias assessments from an included Excel file.



```{r robviz}
rob_file <- system.file("extdata", "beta_carotene_data.xlsx", package = "triangulate")
RoB <- readxl::read_excel(rob_file, sheet = "main_data")
colnames(RoB)[2] <- "Author"

Metadta_RoB_CHD<-merge(Metadta_all_CHD, RoB, by="Author")

# format to robviz
colnames(Metadta_RoB_CHD)[which(colnames(Metadta_RoB_CHD)=="id")]<-"result_id"
colnames(Metadta_RoB_CHD)[which(colnames(Metadta_RoB_CHD)=="Study")]<-"type"
colnames(Metadta_RoB_CHD)[which(colnames(Metadta_RoB_CHD)=="Author")]<-"study"
colnames(Metadta_RoB_CHD)[which(colnames(Metadta_RoB_CHD)=="logRR")]<-"yi"
Metadta_RoB_CHD["vi"]<-Metadta_RoB_CHD["se_logRR"]^2

Metadta_RoB_CHD$study<-as.character(Metadta_RoB_CHD$study)
Metadta_RoB_CHD$type<-as.character(Metadta_RoB_CHD$type)
Metadta_RoB_CHD$Design<-as.character(Metadta_RoB_CHD$Design)

dat_CHDbias<-as_tibble(Metadta_RoB_CHD)

```

### Validate the data frame

explain what check does

```{r}

#check using triangulate
tri_dat_check(dat_CHDbias)

```

### Convert to bias direction format

Convert to long format and map absolute bias directions based on domain judgments.

```{r}

### change domains that doesn't exists in RoB to "None"
dat_CHDbias[is.na(dat_CHDbias)]<-"None"

# convert data to absolute directions of bias/indirectness.
dat_CHDbias_rob <- dat_CHDbias %>%
  tri_to_long() %>%
  tri_absolute_direction() %>%
  tri_to_wide()
```

### Append indirectness

We add indirectness data assuming no major concern across domains.

```{r indirect}
### Indirectness ###
# Creating all as missing

dat_CHDind<-dat_CHDbias %>% select(result_id,study,type,yi,vi)

for (j in paste0("d",1:3)) {
  #j<-"d1"
  dat_CHDind[,paste0(j,"j")] <- rep("None", nrow(dat_CHDind))
  dat_CHDind[,paste0(j,"t")] <- rep("None", nrow(dat_CHDind))
  dat_CHDind[,paste0(j,"d")] <- rep("None", nrow(dat_CHDind))
}

dat_CHDind <- dat_CHDind %>%
  tri_to_long() %>%
  tri_absolute_direction() %>%
  tri_append_indirect(triangulate::dat_ind_values)

```

### Define bias priors and apply bias correction

define what custom priors are doing

```{r bias}
### add critical to prior ###
custom_bias_priors <- triangulate::dat_bias_values %>%
  add_row(domain = 'all',
          j = 'critical',
          bias_m_add = 0.36,
          bias_v_add = 0.2,
          bias_m_prop = 0.12,
          bias_v_prop = 0.064)

dat_CHDbias_prep <- dat_CHDbias %>%
  tri_to_long() %>%
  tri_absolute_direction() %>%
  tri_append_bias(custom_bias_priors)

datCHD_final <- tri_prep_data(dat_CHDbias_prep, dat_CHDind)

```

### Plot bias-adjusted forest plot

```{r plot}

### Plot adjusted with robviz
dat_CHDbias_final <- dat_CHDbias_rob %>%
  right_join(datCHD_final[c("result_id", "yi_adj", "vi_adj")], by = "result_id")

plot_data <- left_join(dat_CHDbias, datCHD_final[, c("result_id", "yi_adj", "vi_adj")])

tri_plot_bias_direction(plot_data)



```

This vignette demonstrated how to use triangulate to conduct a bias-adjusted meta-analysis and visualise estimates for CHD risk and beta-carotene using multiple study designs.

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
